---
title: "Military Expenditure and Power"
subtitle: ""
author: "Jack Cook"
institute: "American University"
date: "05/08/2023"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: left, top

## Project Data

- This dataset includes data on six indicators that make up a state's power: military expenditure, military personnel, energy consumption, iron and steel production, urban population, and total population.
- The data is sourced from the Correlates of War Project. The dataset used is a subset of the original dataset.

- Key variables used include:
  - Year (2000 - 2016)
  - Military Expenditure
  - Military Personnel
  - Country Strength ('Strong' and 'Weak' defined as having a CINC score >.01 or <.01)

Source: *Singer, J. David. 1987. “Reconstructing the Correlates of War Dataset on Material Capabilities of States, 1816-1985” International Interactions, 14: 115-32.*
  
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```


---
class: left, top

## Distribution of Military Personnel Data 2000 - 2016

```{css echo = FALSE}
.left-column {
  float: left;
  width: 35%;
  color: black;
}
.right-column {
  float: right;
  width: 60%;
  height: 100%;
  padding-top: 0;
}
```

```{r, include=FALSE}
  knitr::opts_chunk$set(
    echo = FALSE, dev = 'png', message = FALSE, warning = FALSE, 
    fig.width = 5, fig.asp=0.650, fig.align = 'left', out.width = "100%"
  )
library(tidyverse)
library(xaringan)
library(scales)
library(knitr)
library(wesanderson)
NMC = read.csv('NMC-60-abridged.csv', na = '-9')
NMC2 = subset(NMC, year >= 2000)

```


.right-column[
```{r, fig.retina=2}
options(scipen = 999)

ggplot(NMC2, aes(x = milper)) + 
  geom_histogram(color = 'white', fill = '#E1BD6D') + 
  scale_x_log10(labels = comma) +
  scale_y_continuous(expand = expansion(mult = c(0,0.05))) + 
  labs( title = "Military Personnel by Country 2000 - 2016", y = "Count", 
        x = "Military Personnel (Thousands)" ) + 
  theme_minimal() + 
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major = element_blank()) + 
  geom_vline(xintercept = 22) + 
  annotate("text", x = 62, y = 210, label="Median = 22K", angle = 0)


```
]

.left-column[
- The maximum country military personnel is 2.81 million

- The top 25% of total state military personnel in the data is greater than 84,000

- Median state military personnel is 22,000
]
---
class: left, top

## Military Expenditure: Strong vs. Weak Countries (2000 - 2016)

.right-column[
```{r, fig.retina=2}

library(patchwork)
NMC6 = NMC2 %>% mutate(PowerGroups = ifelse(cinc > .01, 'Strong', 'Weak')) 

regular = ggplot(NMC6, aes(x=year, y=milex)) +
  geom_col(fill = '#0B775E') +
  scale_y_continuous(expand = expansion(mult = c(0,0.05)), 
                     labels = scales::dollar_format()) + 
  labs(title = 'Global Annual Military Expenditure', 
       y = '', x = '') +
  theme_minimal() + 
  theme(panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank())

wrapped = ggplot(NMC6, aes(x=year, y=milex)) +
  geom_col(aes(fill = PowerGroups)) +
  scale_y_continuous(labels = scales::dollar_format()) +
  facet_wrap(~PowerGroups, ncol = 2) + 
  scale_fill_manual(labels = c("Strong", 
                                 "Weak"),
                      values = c("#35274A","#F2300F")) +
  labs(x = "", y = "") +
  theme_minimal() + 
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    strip.text = element_text(face = "bold"))

regular / wrapped

```
]

.left-column[
- Mean Military Expenditure:
  - Weak countries: 1.5 billion USD
  - Strong countries: 54.3 billion USD
  
- In the year 2000, military expenditures totaled 823 billion USD

- In the year 2016, military expenditures totaled 1.5 trillion USD
]

